---
title: "Parking Meter Data Exploration"
author: "Ijeamaka Anyene"
date: "3/15/2020"
output: html_document
---

```{r packages}
library(dplyr)
library(ggplot2)
library(sp)
```

```{r data}
parking_meter_usage_20191101 = readr::read_csv(here::here("data",
"20191101_20200312_SFMTA_Parking_Meter_Detailed_Revenue_Transactions.csv"))
parking_meter_usage_20200313 = readr::read_csv(here::here("data",
"20200313_20200315_SFMTA_Parking_Meter_Detailed_Revenue_Transactions.csv"))

parking_meter_location = readr::read_csv(here::here("data",
                                                    "Parking_Meters.csv"))


# Loading GeoJSON of SF neighborhoods
sf_neighborhoods = rgdal::readOGR(here::here("data", "Analysis Neighborhoods.geojson"))

```

```{r}
parking_meter_usage = rbind(parking_meter_usage_20191101, 
                            parking_meter_usage_20200313)
```

## Q1: How complete is this data? What is the level of missingness?
Are all the POSTIDs within both dataframes?
```{r missing postids}
usage_post_id = unique(parking_meter_usage$POST_ID)
location_post_id = unique(parking_meter_location$POST_ID)  

length(usage_post_id) - sum(usage_post_id %in% location_post_id)
# 212 post IDs in usage DF but not in location DF

length(location_post_id) - sum(location_post_id %in% usage_post_id)
# 6624 post IDs in location DF but not in usage 
```

Are there particular areas where the POSTIDs are missing?
```{r missing maps}
missing_parking_meters = parking_meter_location %>%
  mutate(missing = if_else(POST_ID %in% usage_post_id, "No", "Yes")) %>%
  filter(is.na(LONGITUDE) == FALSE & is.na(LATITUDE) == FALSE)

# Quick Map - There is definitely concentrations of missing parking meters
# in FIDI and possibly where Mission Bay meets SOMA meets the Mission?
ggplot(missing_parking_meters, aes(x = LATITUDE, y = LONGITUDE, colour = missing)) +
  geom_point(size = 0.5)

```

```{r missing longitude}
# Confirmed that this concentration is not due to missing longitude information
sum(is.na(parking_meter_location$LONGITUDE))
```

## Q2: Has there been any high level trends?

```{r joins}
# Joining location and transaction data
colnames_usage = colnames(parking_meter_usage)
parking_meter_combined = left_join(parking_meter_usage, parking_meter_location,
                                   by = c("POST_ID" = "POST_ID")) %>%
  select(all_of(colnames_usage), OBJECTID, JURISDICTION, LONGITUDE, LATITUDE)

# Pulling out date without HMS
parking_meter_combined = parking_meter_combined %>%
  mutate(start_date_new = lubridate::date(SESSION_START_DT))
```

For each day - what are the trends in usage and amount received
```{r}
avg_parking_day = parking_meter_combined %>%
  group_by(start_date_new) %>%
  summarise(avg_gross_paid_amt = mean(GROSS_PAID_AMT, na.rm = TRUE),
            unique_parking_meters = n_distinct(POST_ID))
```

```{r}
ggplot(avg_parking_day, aes(x = start_date_new, y = avg_gross_paid_amt)) +
  geom_line() +
  labs(title = "Average gross paid amount for parking meters per day")

ggplot(avg_parking_day, aes(x = start_date_new, y = unique_parking_meters)) +
  geom_line() +
  labs(title = "Number of unique parking at parking meters per day")
```

SO! There's an interesting trend where on each sunday the cost to park goes up - and less people park. Will remove all sunday dates in the following graphs. Also removing the dates of Christmas, Thanksgiving, and New Years Day where Parking Meters are not enforced. 

```{r}
avg_parking_day = avg_parking_day %>%
  mutate(week_start_date = lubridate::week(start_date_new),
         week_start_date = if_else(week_start_date < 44, week_start_date+53, week_start_date)) %>%
  mutate(day_start_date = lubridate::wday(start_date_new))

```

For each week - what are the trends in usage and amount received?
```{r}
avg_parking_day %>%
  filter(day_start_date  != 1,
         start_date_new != "2019-11-28",
         start_date_new != "2019-12-25",
         start_date_new != "2020-01-01") %>%
  ggplot(., aes(x = start_date_new, y = avg_gross_paid_amt)) +
  geom_line() +
  labs(title = "Average gross paid amount for parking meters per day (excluding Sundays)") +
  geom_smooth(method = "loess")

avg_parking_day %>%
  filter(day_start_date  != 1,
         start_date_new != "2019-11-28",
         start_date_new != "2019-12-25",
         start_date_new != "2020-01-01") %>%
  ggplot(., aes(x = start_date_new, y = unique_parking_meters)) +
  geom_line() +
  labs(title = "Number of unique parking at parking meters per day (excluding Sundays)") +
  geom_smooth(method = "loess")


```
